On the Complexity of Best Arm Identification

### 2017-05-26 Complexity of Best Arm Identification/kl_complexity.Rmd
add theorem 1 of Garivier and Kaufmann

\begin{equation*}
\sum_{a \in S^*_m} \frac{1}{K(\theta_a, \theta_{[m+1]})} + \sum_{a \notin S^*_m \frac{1}{K(\theta_a,\theta_{[m]})}} \leq \mathcal{K}_C(\nu) \leq 24 \min_{\theta \in [\theta_{[m+1]},\theta_{[m]}]} \sum_{a=1}^{K} \frac{1}{K^*(\theta_a, \theta)},
\end{equation*}

@@

>>>>

\begin{equation*}
\sum_{a \in S^*_m} \frac{1}{K(\theta_a, \theta_{[m+1]})} + \sum_{a \notin S^*_m} \frac{1}{K(\theta_a,\theta_{[m]})} \leq \mathcal{K}_C(\nu) \leq 24 \min_{\theta \in [\theta_{[m+1]},\theta_{[m]}]} \sum_{a=1}^{K} \frac{1}{K^*(\theta_a, \theta)},
\end{equation*}

@@

### Optimal Best Arm Identification

A very similar application of the KL divergence has been given in Garivier and Kaufmann to derive a lower bound on the sample complexity in the case of best arm identification. In particular, the transportation Lemma 1 is used again. And a matching algorithm for $K \geq 2$ arms is proposed.

Talk about:
- Not a simple sum of squares
- Idea of algorithm: Equalize the probability of all wrong decisions

In Garivier and Kaufmann (2016), the authors are able to derive a somewhat similar lower bound on the sample complexity of a $K$-armed bandit in the single best arm identification problem with fixed confidence. Additionally, they provide a $\delta$-PAC algorithm achieving said bound asymptotically.

In particular, the authors pick up Lemma 1 from Kaufmann et al. (2016) -- which we discussed above -- to derive the lower bound. While presenting these results, we try to stick as much as possible with the notation used so far. In particular, we let for distributions $\nu_{\theta}$ from a canonical exponential family, which can be parameterized by its mean $\stackrel{.}{b}(\theta) = \mu$ and thus be written as $\nu^{\mu}$; for these, we continue to write the Kullback-Leibler divergence as

\begin{equation*}
\KL(\mu, \mu') = \KL(\nu^{\mu},\nu^{\mu'}) = \KL(\nu_{\theta},\nu_{\theta'}) = \KL(\nu, \nu')
\end{equation*}

while $d(x,y)$ continues to represent the binary relative entropy.

As we will consider only problems with exponential family distributions, the distributions can be parameterized by their means. Therefore, if we let $\mathcal{S}$ to be a set of bandit models with $K$ distributions each, then each model is fully defined by a vector $\bm{\mu} = (\mu_1, \dots, \mu_K)$. Assume that every $\bm{\mu} \in \mathcal{S}$ has a unique maximum given by $a^*(\bm{\mu})$ such that $\mu_{a^*} > \max_{a \neq a^*(\bm{\mu})} \mu_a$. In order to fix the confidence, we specify the risk $\delta$. The strategies corresponding to this risk are called $\delta$-PAC if for every $\bm{\mu} \in \mathcal{S}$, $\mathbb{P}_{\bm{\mu}}(\tau_{\delta} < \infty) = 1$ (the strategy ends surely after a finite number of samples) and $\mathbb{P}_{\bm{\mu}}(\hat{a}_{\tau_{\delta}} \neq a^*) \leq \delta$ (the probability of picking not the optimal arm is smaller than the fixed risk). As in Garivier and Kaufmann (2016), we introduce the set of problems for which the optimal arm is different from the one in $\bm{\mu}$:

\begin{equation*}
\text{Alt}(\bm{\mu}) := \{\bm{\lambda} \in \mathcal{S}: a^*(\bm{\lambda}) \neq a^*(\bm{\mu})\}
\end{equation*}

where $\bm{\lambda}$ is the vector of means defining the distributions of the alternative problem to $\bm{\mu}$. Lastly, we have the set of probability distributions on the arms $\mathcal{A}$: $\Sigma_K = \{w \in \mathbb{R}_{+}^K: w_1 + \dots + w_K = 1\}$. \marginpar{Give more intuition for what these probability distributions are} We are now ready to state the theorem.

\begin{theorem}[Theorem 1, Garivier and Kaufmann, 2016] \label{theorem:GarivierKaufmannTheorem1}
Let $\delta \in (0,1)$. For any $\delta$-PAC strategy and any bandit model $\bm{\mu} \in \mathcal{S}$,

\begin{equation*}
\mathbb{E}_{\bm{\mu}} [\tau_{\delta}] \geq T^*(\bm{\mu}) \text{\em{d}}(\delta,1-\delta),
\end{equation*}

where

\begin{equation} \label{eq:GarivierKaufmann_Theorem1_Game}
T^*(\bm{\mu})^{-1} := \sup_{w \in \Sigma_K} \inf_{\bm{\lambda} \in \text{\em{Alt}}(\bm{\mu})} \left( \sum_{a=1}^K w_a \text{\em{KL}}(\mu_a, \lambda_a)\right) .
\end{equation}
\end{theorem}

And so we see that the two bounds on the sample complexity (one for Top-$m$, one for best arm identification) differ only in the way that they build the sum of Kullback-Leibler divergences over the $K$ arms. In contrast to Theorem \ref{theorem:KaufmannEtAl2016Theorem1}, here the contribution of each arm through its KL-divergence is weighted. The difference is that while we were able to very clearly describe the change of distribution in the Top-$m$ case for each arm -- for each arm, we specified an alternative model that yields a new set of $m$ optimal arms -- we now consider all alternative models that yield a different best arm than the model $\bm{\mu}$. In the previous theorem, we sum over the complexity of each individual arm, as the change in distribution only ever touches one arm, thus leading to a KL-divergence of 0 for the other arms while still changing the optimal set. Here, we consider all instances of alternative models $\bm{\lambda}$, and thus also those in which several arms are modified (and thus have a KL-divergence unequal 0 in the result of Lemma 1).

*Proof of Theorem \ref{theorem:GarivierKaufmannTheorem1}*: As stated before, we let $\delta \in (0,1)$, $\bm{\mu} \in \mathcal{S}$, and only consider $\delta$-PAC strategies. As before, $N_a(t)$ describes the random number of pulls of arm $a$ until time $t$, for all $t\geq 1$. Again, Lemma \ref{theorem:KaufmannEtAlLemma1} states how the failure probability $\delta$ relates to the expected number of draws of each arm given two models which differ in their optimal arm, and where it is evaluated based on the Kullback-Leibler divergence for each pair of arms how much the two models differ:

\begin{equation*}
\forall \lambda \in \mathcal{S}: a^*(\lambda) \neq a^*(\mu), \sum_{a=1}^K \KL(\mu_a,\lambda_a)\mathbb{E}_{\bm{\mu}}[N_a(\tau_{\delta})] \geq \d(\delta, 1- \delta).
\end{equation*}

As mentioned, we now do not consider a specific alternative for each arm, but instead look at the infimum over all the possible alternatives, and use the fact that $\mathbb{E}_{\bm{\mu}}[\tau_{\delta}] = \sum_{a=1}^K \mathbb{E}_{\bm{\mu}}[N_a(\tau_{\delta})]$:

\begin{align*}
\der(\delta, 1-\delta) & \leq \inf_{\lambda \in \text{Alt}(\bm{\mu})} \mathbb{E}_{\bm{\mu}}[\tau_{\delta}] \Big(\sum_{a=1}^K \frac{\mathbb{E}_{\bm{\mu}}[N_a]}{\mathbb{E}_{\mu}[\tau_{\delta}]} \KL(\mu_a, \lambda_a)\Big) \\
& = \mathbb{E}_{\bm{\mu}}[\tau_{\delta}] \inf_{\lambda \in \text{Alt}(\bm{\mu})} \Big(\sum_{a=1}^K \frac{\mathbb{E}_{\bm{\mu}}[N_a]}{\mathbb{E}_{\mu}[\tau_{\delta}]} \KL(\mu_a, \lambda_a)\Big) \\
& \leq \mathbb{E}_{\bm{\mu}}[\tau_{\delta}] \sup_{w \in \Sigma_K} \inf_{\lambda \in \text{Alt}(\bm{\mu})} \Big(\sum_{a=1}^K w_a \KL(\mu_a, \lambda_a)\Big).
\end{align*}

As stated in Garivier and Kaufmann (2016), the term $\frac{\mathbb{E}_{\bm{\mu}}[N_a]}{\mathbb{E}_{\mu}[\tau_{\delta}]}$ gives the fraction of draws each arm receives. This value, however, depends on the used algorithm. To make it hold for any $\delta$-PAC algorithm, it is replaced by the supremum over $\Sigma_K$, which is the distribution over these fractions for any $\mathcal{A}$, that is, for any $\delta$-PAC strategy.

This lower bound can be interpreted as a Minimax game, where the experimenter has to choose the share of samples he attributes to each arm in a way that is optimal given that nature will respond and choose the most difficult alternative model given his weights. Thus, to make the lower bound practical, it is a matter of finding the optimal weights $w^*(\bm{\mu})$.

### 2017-05-26 Complexity of Best Arm Identification/kl_complexity.tex
add theorem 1 of Garivier and Kaufmann

>>>>

\subsubsection{Optimal Best Arm
Identification}\label{optimal-best-arm-identification}

A very similar application of the KL divergence has been given in
Garivier and Kaufmann to derive a lower bound on the sample complexity
in the case of best arm identification. In particular, the
transportation Lemma 1 is used again. And a matching algorithm for
\(K \geq 2\) arms is proposed.

Talk about: - Not a simple sum of squares - Idea of algorithm: Equalize
the probability of all wrong decisions

In Garivier and Kaufmann (2016), the authors are able to derive a
somewhat similar lower bound on the sample complexity of a \(K\)-armed
bandit in the single best arm identification problem with fixed
confidence. Additionally, they provide a \(\delta\)-PAC algorithm
achieving said bound asymptotically.

In particular, the authors pick up Lemma 1 from Kaufmann et al. (2016)
-- which we discussed above -- to derive the lower bound. While
presenting these results, we try to stick as much as possible with the
notation used so far. In particular, we let for distributions
\(\nu_{\theta}\) from a canonical exponential family, which can be
parameterized by its mean \(\stackrel{.}{b}(\theta) = \mu\) and thus be
written as \(\nu^{\mu}\); for these, we continue to write the
Kullback-Leibler divergence as

\begin{equation*}
\KL(\mu, \mu') = \KL(\nu^{\mu},\nu^{\mu'}) = \KL(\nu_{\theta},\nu_{\theta'}) = \KL(\nu, \nu')
\end{equation*}

while \(d(x,y)\) continues to represent the binary relative entropy.

As we will consider only problems with exponential family distributions,
the distributions can be parameterized by their means. Therefore, if we
let \(\mathcal{S}\) to be a set of bandit models with \(K\)
distributions each, then each model is fully defined by a vector
\(\bm{\mu} = (\mu_1, \dots, \mu_K)\). Assume that every
\(\bm{\mu} \in \mathcal{S}\) has a unique maximum given by
\(a^*(\bm{\mu})\) such that
\(\mu_{a^*} > \max_{a \neq a^*(\bm{\mu})} \mu_a\). In order to fix the
confidence, we specify the risk \(\delta\). The strategies corresponding
to this risk are called \(\delta\)-PAC if for every
\(\bm{\mu} \in \mathcal{S}\),
\(\mathbb{P}_{\bm{\mu}}(\tau_{\delta} < \infty) = 1\) (the strategy ends
surely after a finite number of samples) and
\(\mathbb{P}_{\bm{\mu}}(\hat{a}_{\tau_{\delta}} \neq a^*) \leq \delta\)
(the probability of picking not the optimal arm is smaller than the
fixed risk). As in Garivier and Kaufmann (2016), we introduce the set of
problems for which the optimal arm is different from the one in
\(\bm{\mu}\):

\begin{equation*}
\text{Alt}(\bm{\mu}) := \{\bm{\lambda} \in \mathcal{S}: a^*(\bm{\lambda}) \neq a^*(\bm{\mu})\}
\end{equation*}

where \(\bm{\lambda}\) is the vector of means defining the distributions
of the alternative problem to \(\bm{\mu}\). Lastly, we have the set of
probability distributions on the arms \(\mathcal{A}\):
\(\Sigma_K = \{w \in \mathbb{R}_{+}^K: w_1 + \dots + w_K = 1\}\).
\marginpar{Give more intuition for what these probability distributions are}
We are now ready to state the theorem.

\begin{theorem}[Theorem 1, Garivier and Kaufmann, 2016] \label{theorem:GarivierKaufmannTheorem1}
Let $\delta \in (0,1)$. For any $\delta$-PAC strategy and any bandit model $\bm{\mu} \in \mathcal{S}$,

\begin{equation*}
\mathbb{E}_{\bm{\mu}} [\tau_{\delta}] \geq T^*(\bm{\mu}) \text{\em{d}}(\delta,1-\delta),
\end{equation*}

where

\begin{equation} \label{eq:GarivierKaufmann_Theorem1_Game}
T^*(\bm{\mu})^{-1} := \sup_{w \in \Sigma_K} \inf_{\bm{\lambda} \in \text{\em{Alt}}(\bm{\mu})} \left( \sum_{a=1}^K w_a \text{\em{KL}}(\mu_a, \lambda_a)\right) .
\end{equation}
\end{theorem}

And so we see that the two bounds on the sample complexity (one for
Top-\(m\), one for best arm identification) differ only in the way that
they build the sum of Kullback-Leibler divergences over the \(K\) arms.
In contrast to Theorem \ref{theorem:KaufmannEtAl2016Theorem1}, here the
contribution of each arm through its KL-divergence is weighted. The
difference is that while we were able to very clearly describe the
change of distribution in the Top-\(m\) case for each arm -- for each
arm, we specified an alternative model that yields a new set of \(m\)
optimal arms -- we now consider all alternative models that yield a
different best arm than the model \(\bm{\mu}\). In the previous theorem,
we sum over the complexity of each individual arm, as the change in
distribution only ever touches one arm, thus leading to a KL-divergence
of 0 for the other arms while still changing the optimal set. Here, we
consider all instances of alternative models \(\bm{\lambda}\), and thus
also those in which several arms are modified (and thus have a
KL-divergence unequal 0 in the result of Lemma 1).

\emph{Proof of Theorem \ref{theorem:GarivierKaufmannTheorem1}}: As
stated before, we let \(\delta \in (0,1)\),
\(\bm{\mu} \in \mathcal{S}\), and only consider \(\delta\)-PAC
strategies. As before, \(N_a(t)\) describes the random number of pulls
of arm \(a\) until time \(t\), for all \(t\geq 1\). Again, Lemma
\ref{theorem:KaufmannEtAlLemma1} states how the failure probability
\(\delta\) relates to the expected number of draws of each arm given two
models which differ in their optimal arm, and where it is evaluated
based on the Kullback-Leibler divergence for each pair of arms how much
the two models differ:

\begin{equation*}
\forall \lambda \in \mathcal{S}: a^*(\lambda) \neq a^*(\mu), \sum_{a=1}^K \KL(\mu_a,\lambda_a)\mathbb{E}_{\bm{\mu}}[N_a(\tau_{\delta})] \geq \d(\delta, 1- \delta).
\end{equation*}

As mentioned, we now do not consider a specific alternative for each
arm, but instead look at the infimum over all the possible alternatives,
and use the fact that
\(\mathbb{E}_{\bm{\mu}}[\tau_{\delta}] = \sum_{a=1}^K \mathbb{E}_{\bm{\mu}}[N_a(\tau_{\delta})]\):

\begin{align*}
\der(\delta, 1-\delta) & \leq \inf_{\lambda \in \text{Alt}(\bm{\mu})} \mathbb{E}_{\bm{\mu}}[\tau_{\delta}] \Big(\sum_{a=1}^K \frac{\mathbb{E}_{\bm{\mu}}[N_a]}{\mathbb{E}_{\mu}[\tau_{\delta}]} \KL(\mu_a, \lambda_a)\Big) \\
& = \mathbb{E}_{\bm{\mu}}[\tau_{\delta}] \inf_{\lambda \in \text{Alt}(\bm{\mu})} \Big(\sum_{a=1}^K \frac{\mathbb{E}_{\bm{\mu}}[N_a]}{\mathbb{E}_{\mu}[\tau_{\delta}]} \KL(\mu_a, \lambda_a)\Big) \\
& \leq \mathbb{E}_{\bm{\mu}}[\tau_{\delta}] \sup_{w \in \Sigma_K} \inf_{\lambda \in \text{Alt}(\bm{\mu})} \Big(\sum_{a=1}^K w_a \KL(\mu_a, \lambda_a)\Big).
\end{align*}

As stated in Garivier and Kaufmann (2016), the term
\(\frac{\mathbb{E}_{\bm{\mu}}[N_a]}{\mathbb{E}_{\mu}[\tau_{\delta}]}\)
gives the fraction of draws each arm receives. This value, however,
depends on the used algorithm. To make it hold for any \(\delta\)-PAC
algorithm, it is replaced by the supremum over \(\Sigma_K\), which is
the distribution over these fractions for any \(\mathcal{A}\), that is,
for any \(\delta\)-PAC strategy.

This lower bound can be interpreted as a Minimax game, where the
experimenter has to choose the share of samples he attributes to each
arm in a way that is optimal given that nature will respond and choose
the most difficult alternative model given his weights. Thus, to make
the lower bound practical, it is a matter of finding the optimal weights
\(w^*(\bm{\mu})\).
